\chapter*{Introduction}
\label{cha:intro}

This project has two goals. First we want to check whether a proof
assistant can do differential topology. Many people still think that
formal mathematics are mostly suitable for algebra, combinatorics, or
foundational studies. So we chose one of the most famous examples
of geometric topology theorems associated to tricky geometric intuition:
the existence of sphere eversions. Note however that we won't focus on
any of the many videos of explicit sphere eversions. We will prove a
general theorem which immediately implies the existence of sphere
eversions.

The second goal of this project is to experiment using a formalization
blueprint that evolves with the project until we get a proof that has
very closely related formal and informal presentations.

In this introduction, we will describe the mathematical context of this
project, the main definitions and statements, and outline the proof
strategy.

Gromov observed that it's often fruitful to distinguish two kinds of geometric
construction problems. He says that a geometric construction problem satisfies
the $h$-principle if the only obstructions to the existence of a solution come
from algebraic topology. In this case, the construction is called flexible,
otherwise it is called rigid. This definition is purposely vague. We
will see a rather general way to give it a precise meaning, but one must keep
in mind that such a precise meaning will fail to encompass a number of
situations that can be illuminated by the $h$-principle dichotomy point of view.

The easiest example of a flexible construction problem which is not totally
trivial and is algebraically obstructed is the deformation of immersions
of circles into planes. Let $f_0$ and $f_1$ be two maps from $𝕊^1$ to $ℝ^2$
that are immersions. Since $𝕊^1$ has dimension one, this mean that both
derivatives $f_0'$ and $f_1'$ are nowhere vanishing maps from $𝕊^1$ to
$ℝ^2$. The geometric object we want to construct is a (smooth) homotopy
of immersions from $f_0$ to $f_1$, ie a smooth map $F \co 𝕊^1 × [0, 1] → ℝ^2$
such that $\rst{F}{𝕊^1 × \{0\}} = f_0$, $\rst{F}{𝕊^1 × \{1\}} = f_1$, and
each $f_p := \rst{F}{𝕊^1 × \{p\}}$ is an immersion. If such a homotopy
exists then, $(t, p) ↦ f_p'(t)$ is a homotopy from $f_0'$ to $f_1'$ among
maps from $𝕊^1$ to $ℝ^2 ∖ \{0\}$. Such maps have a well defined winding
number $w(f'_i) ∈ ℤ$ around the origin, the degree of the normalized map
$f'_i/‖f'_i‖ \co 𝕊^1 → 𝕊^1$. So $w(f_0') = w(f_1')$ is a necessary condition
for the existence of $F$, which comes from algebraic topology.  The
Whitney--Graustein theorem states that this necessary condition is also
sufficient. Hence this geometric construction problem is flexible.  One
can give a direct proof of this result, but it will also
follow from general results proved in this project.

An important lesson from the above example is that algebraic topology
can give us more than a necessary condition. Indeed the (one-dimensional)
Hopf degree theorem ensures that, provided $w(f'_0) = w(f'_1)$, there
exists a homotopy $g_p$ of nowhere vanishing maps relating $f'_0$ and
$f'_1$. We also know from the topology of $ℝ^2$ that $f_0$ and $f_1$ are
homotopic, say using the straight-line homotopy
$p ↦ f_p = (1-p)f_0 + pf_0$. But there is no a priori relation
between $g_p$ and the derivative of $f_p$ for $p ∉ \{0, 1\}$.
So we can restate the crucial part of the
Whitney--Graustein theorem as: there is a homotopy of immersion from
$f_0$ to $f_1$ as soon as there is (a homotopy from $f_0$ to $f_1$) and
a homotopy from $f_0'$ to $f_1'$ among nowhere vanishing maps. The
parenthesis in the previous sentence indicated that this condition is
always satisfied, but it is important to keep in mind for
generalizations. Gromov says that such a homotopy of uncoupled pairs
$(f, g)$ is a formal solution of the original problem.

One can generalize this discussion of uncoupled maps replacing a
map and its derivative. This is pretty easy for maps from a manifold $M$
to a manifold $N$. The so called $1$-jet space $J^1(M, N)$ is the
space of triples $(m, n, φ)$ with $m ∈ M$, $n ∈ N$, and
$φ ∈ \Hom(T_mM, T_nN)$, the space of linear maps from $T_mM$ to $T_nN$.
One can define a smooth manifold structure on $J^1(M, N)$, of dimension
$\dim(M) + \dim(N) + \dim(M)\dim(N)$ which fibers over $M$, $N$ and
their product $J^0(M, N) := M × N$. Beware that the notation $(m, n, φ)$
does not mean that $J^1(M, N)$ is a product of three manifolds, the space
where $φ$ lives depends on $m$ and $n$.
Any smooth map $f \co M → N$ gives rise to a section $j^1f$ of
$J^1(M, N) → M$ defined by $j^1f(m) = (m, f(m), T_m f)$. Such a section is
called a \emph{holonomic section} of $J^1(M, N)$.
In the Whitney--Graustein example, we use the canonical trivialization of
$T𝕊^1$ and $Tℝ^2$ to represent $j^1f$ has a pair of maps $(f, f')$. The
role played by $(f, g)$ in this example is played in general by sections
of $J^1(M, N) → M$ which are not necessarily holonomic.

One can generalize this discussion to $J^r(M, N)$ which remembers
derivatives of maps up to order $r$ for some given $r ≥ 0$. One can also
consider sections of an arbitrary bundle $E → M$ instead of functions
from $M$ to $N$, which are sections of the trivial bundle $M × N → N$.
But the case of $J^1(M, N)$ will be sufficient for this project.

\begin{definition-intro}
  A first order differential relation $\Rel$ for maps from $M$ to
  $N$ is a subset of $J^1(M, N)$. A solution of $\Rel$ is a function
  $f \co M \to N$ such that $j^1f(m)$ is in $\Rel$ for all $m$.
  A \emph{formal solution} of $\Rel$ is a non-necessarily holonomic
  section of $J^1(M, N) → M$ which takes value in $\Rel$.

  The partial differential relation $\Rel$ satisfies the $h$-principle
  if any formal solution $σ$ of $\Rel$ is homotopic, among formal
  solutions, to some holonomic one $j^1f$.
\end{definition-intro}

For instance, an immersion of $M$ into $N$ is a solution of
\[
\Rel = \{(m, n, φ) ∈ J^1(M, N) \;|\; φ \text{ is injective}\}.
\]
As we saw with
the Whitney--Graustein problem, we are not only interested to individual
solutions, but also in families of solutions. In differential
topology, a smooth family of maps between manifolds $X$ and $Y$ is a smooth map
$h \co P × X → Y$ seen as the collection of maps $h_p \co x ↦ h(p, x)$. Here $P$
stands for ``parameter space''. A smooth family of sections of $E → X$ is a
smooth family of maps $σ \co P × X → E$ such that each $σ_p$ is a section.

When the parameter space $P$ has boundary, we will typically assume that formal
solutions $σ_p$ are holonomic for $p$ in $𝓝(∂P)$. This is an abreviation
meaning: ``there is an unspecified neighborhood $U$ of $∂P$ such that
$σ_p$ is holonomic for $p$ in $U$''. Note that an unspecified neighborhood can
change from invocation to invocation. For instance in the next
definition, the second unspecified neighborhood is typically smaller
than the first one.


\begin{definition-intro}
	A partial differential relation $\Rel$ satisfies the parametric
  $h$-principle if every family of formal solutions
  $σ \co M × P → J^1(M, N)$ which are holonomic for $p$ in $𝓝(∂P)$ is
  homotopic, relative to $𝓝(∂P)$, to a family of holonomic sections.
\end{definition-intro}

There are other variations on this definition. For instance a formal solution
could be holonomic on $𝓝(A)$ for some subset $A$ of $M$, and we say that $\Rel$
satisfies the relative $h$-principle if $σ$ can be deformed to a holonomic
solution without changing it on $𝓝(A)$.

One can also insist on the deformed solution to be $C^0$-close to the original
one. In this case one talks about a $C^0$-dense $h$-principle.
We are now ready to state our main goal.

\begin{theorem-intro}
	The relation of immersions in positive codimension (ie immersions of $M$
	into $N$ with $\dim(N) > \dim(M)$) satisfies all forms of $h$-principles.
\end{theorem-intro}

This theorem covers the Whitney--Graustein theorem (in its second
form, assuming the existence of a homotopy between derivatives). But there are
much less intuitive applications. The most famous one is the existence of
sphere eversions: one can ``turn $𝕊^2$ inside-out among immersions of $𝕊^2$ into
$ℝ^3$).

\begin{corollary-intro}[Smale 1958]
	There is a homotopy of immersion of $𝕊^2$ into $ℝ^3$ from the inclusion map to
	the antipodal map $a \co q ↦ -q$.
\end{corollary-intro}

The reason why this is turning the sphere inside-out is that $a$ extends as
a map from $ℝ^3 ∖ \{0\} → ℝ^3 ∖ \{0\}$ by
\[
	\hat{a} \co q ↦ -\frac{1}{‖q‖^2}q
\]
which exchanges the interior and exterior of $𝕊^2$. More abstractly, one can say
the normal bundle of $𝕊^2$ is trivial, hence one can extend $a$ to a tubular
neighborhood of $𝕊^2$ as an orientation preserving map. Since $a$ is orientation
reversing, any such extension will be reversing coorientation.

\begin{proof}[Proof of the sphere eversion corollary]
	We denote by $ι$ the inclusion of $𝕊^2$ into $ℝ^3$.
	We set $j_t = (1-t)ι	+ ta$. This is a homotopy from $ι$ to $a$ (but not an
	immersion for $t=1/2$).
        We need to check there is no obstruction to building a homotopy of
        formal solutions above those maps.
        One could show that the relevant homotopy group (replacing $π_1(𝕊^1)$
        from the Whitney--Graustein example) is $\pi_2(\SO_3(ℝ))$.
        This group is trivial, hence there is no obstruction.
        But actually we can write an explicit homotopy here, without using any
        algebraic topology.
        Using the canonical trivialization of the tangent
	bundle of $ℝ^3$, we can set, for $(q, v) ∈ T𝕊^2$,
	$G_t(q, v) = \mathrm{Rot}_{Oq}^{πt}(v)$, the rotation around axis $Oq$ with
	angle $πt$. The family $σ \co t ↦ (j_t, G_t)$ is a homotopy of formal immersions
	relating $j^1ι$ to $j^1a$. The above theorem ensures this family is homotopic,
	relative to $t = 0$ and $t = 1$, to a family of holonomic formal immersions,
	ie a family $t ↦ j^1f_t$ with $f_0 = ι$, $f_1 = a$, and each $f_t$ is an
	immersion.
\end{proof}

The theorem above follows from a more general theorem which is slightly too
technical for this introduction: the $h$-principle for open and ample
first order differential relations. We will prove this theorem using a
technique which is even more general: convex integration. For instance
this technique also underlies the constructions of paradoxical isometric
embeddings, which could be a nice follow-up project.

We'll end this introduction by describing the key construction of convex
integration, since it is very nice and elementary. Convex integration
was invented by Gromov around 1970, inspired in particular by the
$C^1$ isometric embedding work of Nash and the original proof of
flexibility of immersions. This term is pretty vague however, and there
are several different implementations. The newest one, and by far the
most efficient one, is Mélanie Theillière's corrugation process from
2017. And this is what we will use.

Let $f$ be a map from $ℝ^n$ to $ℝ^m$. Say we want to turn $f$ into a
solution of some partial differential relation. For instance if we are
interested in immersions, we want to make sure its differential (or
equivalently its Jacobian matrix) is everywhere injective. We will ensure
this by tackling each partial derivative in turn. In the immersion
example, we first make sure $∂_1f(x) := ∂f(x)/∂x_1$ is non-zero for all $x$.
Then we make sure $∂_2f(x)$ is not colinear to
$∂_1f(x)$. Then we make sure $∂_3f(x)$ is not in the
plane spanned by the two previous derivatives, etc\dots\ until
all $n$ partial derivatives are everywhere linearly independent.

In general, what happens is that, for each number $j$ between $1$ and
$n$, we wish $∂_jf(x)$ could live in some open subset
$Ω_x ⊂ ℝ^m$. Assume there is a
smooth compactly supported family of loops $γ \co ℝ^n × 𝕊^1 → ℝ^m$ such
that each $γ_x$ takes values in $Ω_x$, and has average value
$\int_{𝕊^1} γ_x = ∂_j f(x)$. Obviously such loops
can exist only if $∂_jf(x)$ is in the convex hull of $Ω_x$,
hence the name convex integration, and we will see this condition is
almost sufficient. In the immersion case, this convex hull condition
will always be met because, from the above description, we see that
$Ω_x$ will always be the complement of a linear subspace with
codimension at least two.

For some large positive $N$, we replace $f$ by the new map
\[
  x ↦ f(x) + \frac1N  ∫_0^{Nx_j} \left[γ_x(s) - ∂_jf(x)\right]ds.
\]

A wonderfully easy exercise shows that, provided $N$ is large enough,
we have achieved $∂_j f(x) ∈ Ω_x$, almost without modifying
derivatives $∂_i f(x)$ for $i ≠ j$, and almost without moving
$f(x)$.

In addition, if we assume that $γ_x$ is constant (necessarily with value
$∂_j f(x)$) for $x$ near some subset $K$ where
$∂_j f(x)$ was already good, then nothing changed on $K$ since the
integrand vanishes there. It is also easy to damp out this modification
by multiplying the integral by a cut-off function. So this is a very
local construction, and it isn't obvious how the absence of homotopical
obstruction, embodied by the existence of a formal solution, should
enter the discussion. The answer is that is essentially provides a way
to coherently choose base points for the $γ_x$ loops.

\Cref{chap:loops} provides the loops supply. \Cref{chap:local} then
discusses the local theory, including the key construction above.
It also proves the sphere eversion corollary. The full project, also
includes a chapter about the global theory.

